\begin{thebibliography}{99}

\bibitem{Barták1998} Barták, Roman (1998) Meta-Interpreters. in \textit{Online Prolog Programming} (http://kti.ms.mff.cuni.cz/~bartak/prolog/meta\_interpret.html)

\bibitem{berger2004}
Berger, Shai (2004) PYTHOLOGIC -- PROLOG SYNTAX IN PYTHON (PYTHON RECIPE)

\bibitem{Bolz2007} Bolz, Carl Friedrich (2007) A Prolog Interpreter in Python.

\bibitem{Cesar2019} Cesar, Bruno Kim Medeiros (2019) Prol: a minimal, inefficient Prolog interpreter in a few LOCs of Python (https://gist.github.com/brunokim/0a737a8642b756a5d0dcc3a07ec1ef81)

\bibitem{Delford2009} Delord, Christophe (2009) PyLog

\bibitem{Frederiksen2011} Frederiksen, Bruce (2011) Pyke

\bibitem{Kopec2019} Constraint-Satisfaction Problems in Python. Chapter 3 of Kopec (2019) \textit{Classic Computer Science Problems in Python}

\bibitem{Maxime2016} Maxime, Istasse (2016) Prology: Logic programming for Python3

\bibitem{Meyers2015} Prolog in Python.

\bibitem{Miljkovic2019} 
Miljkovic, Nikola (2019) Python Prolog Interpreter.
% \href{https://github.com/photonlines/Python-Prolog-Interpreter}{Python Prolog Interpreter}. 

\bibitem{Niemeyer2019} Niemeyer, Gustavo and Sébastien Celles (2019). python-constraint library. Pypi page: python-constraint (Also discussed in Popović, Olivera (2019) Constraint Programming with python-constraint) 

\bibitem{Piumarta2017} Piumarta, Ian (2017), Lecture notes and slides from weeks 5-7 of a course on programming paradigms. % http://www.ritsumei.ac.jp/~piumarta/pl/

\bibitem{Raturi2019}  Raturi, Gaurav (2019) 5 Best Programming Languages to choose for Developing Innovative AI Solutions, \textit{Becoming Human}.

\bibitem{Rocklin2019} Rocklin, Matthew (2019) kanren: Logic Programming in Python

\bibitem{Sagar2019} Sagar, Paresh  (2019) Powerful Programming Languages to Code your AI Application. \textit{medium.com}.

\bibitem{Santini2018} Santini, Claudio (2018) Pampy: The Pattern Matching for Python you always dreamed of.

\bibitem{Thompson2017} Thompson, Jeff (2017) Yield Prolog

\bibitem{Tubella1994} Tubella, J., \& González, A. (1994). Combining depth-first and breadth-first search in Prolog execution. In GULP-PRODE (2)(pp. 452-453)

\bibitem{Triska2016}
Triska, Markus (2016) The {Boolean} Constraint Solver of SWI-Prolog. in Kiselyov, Oleg and Andy King (eds.) \textit{Functional and Logic Programming. 13th International Symposium}, Springer (LNCS) 45--61.  
% \href{http://eu.swi-prolog.org/man/clpb.html}{CLP(FD): Constraint Logic Programming over Finite Domains}

\begin{comment}
This is a more proper representation of the preceding--although it doesn't list Springer as the publisher. Can we get LaTeX to handle it?
@inproceedings{Triska2016,
  author    = "Markus Triska",
  title     = "The {Boolean} Constraint Solver of {SWI-Prolog}:
               System Description",
  booktitle = "FLOPS: Functional and Logic Programming. 13th International Symposium",
  series    = "LNCS",
  volume    = 9613,
  year      = 2016,
  pages     = "45--61"
}
\end{comment}


\bibitem{Triska2019}
Triska, Markus (2019) \textit{The Power of Prolog} 

% \bibitem{Weinberger1974} Weinberger, D. B. (1974). Sufficient regularity conditions for common transversals. Journal of Combinatorial Theory, Series A, 16(3), 380-390.

% \bibitem{CryptPuzzleSolver} http://bach.istc.kobe-u.ac.jp/llp/crypt.html

% \bibitem{BritannicaCrypt} https://www.britannica.com/science/cryptarithm


\end{thebibliography}
